---
title: "Chapter_1_Data_Visualization"
author: "Rashad Long"
date: "`r Sys.Date()`"
output: html_document
warnings: false
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```


### Chapter 1 of R for Data Science

```{r penguins}
penguins
```

### alternative view, where you can see all variables and the first few observations of each variable

```{r glimpse-view}
glimpse(penguins)
View(penguins)
```

### To learn more about penguins

```{r penguins-help}
?penguins
```

### Our ultimate goal in this chapter is to recreate the following visualization displaying the relationship between flipper lengths and body masses of these penguins, taking into consideration the species of the penguin.

```{r penguins-plot}
ggplot(data=penguins,
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) + geom_point(mapping = aes(color=species, shape = species))+ geom_smooth(method = "lm") + labs( title = "Body mass and flipper length", subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo penguins", x = "Flipper length (mm)", y = "Body mass (g)", color = "Species", shape = "Species") + scale_color_colorblind()
```

1.  How many rows are in penguins? How many columns?

```{r penguins-rows-columns}
nrow(penguins)
ncol(penguins)
```
2.    What does the bill_depth_mm variable in the penguins data frame describe? Read the help for ?penguins to find out.
```{r penguins-bill-depth}
?penguins

```

3.   Make a scatterplot of bill_depth_mm vs. bill_length_mm. That is, make a scatterplot with bill_depth_mm on the y-axis and bill_length_mm on the x-axis. Describe the relationship between these two variables
```{r penguins-plot2}
ggplot(data=penguins, mapping=aes(x = bill_length_mm, y =bill_depth_mm)) + geom_point()
```
















