---
title: "String_Manip_&_Date_Processing"
author: "TT"
date: "9/14/2019"
output: html_document
---

## A. Working with Strings

### Regular expressions

```{r}
fruits <- c('Apple', 'Banana', 'Orange', 'Grape', 'Pineapple', 'Kiwi', 'Peach', 'Mango', 'Strawberry', 'Guava', 'Cherry', 'Apple', 'banana')

fruits =="Banana" #exact match
```

```{r}
which(fruits == "Banana") #use which to get position
```

```{r}
breakfast = c("Apple", "Banana", "Apple", "banana")
fruits %in% breakfast 
```

```{r}
match(breakfast, fruits) #notice that only the first match is returned
```

```{r}
fruits == "bana" #how do we search for pattern?
```

*grep:* Identifying strings that match a particular criteria
```{r}
grep(pattern = 'bana', fruits, value = TRUE, ignore.case = TRUE)
```

*gsub:* Replacing instances of a string with another of your choosing
```{r}
gsub(pattern = "Ch", replacement = "B", fruits[11])
```

```{r}
breakfast.ab = c("App", "bana") 
pmatch(breakfast.ab, fruits) #notice that Apple is not unique so it won't work
```

```{r}
grep("nana", fruits) #grep works but one pattern at a time
```

*gregexpr:* Identifying positions of a string of interest
```{r}
(positions_a <- gregexpr(pattern = "a", text = fruits, ignore.case = TRUE))
```
positions_a
### Paste

Use seperators to combine strings with other characters or numbers
```{r}
paste("X", 1:5, sep = ".")
```

Use collapse to combine multiple string outputs together
```{r}
paste("X", 1:5, sep = ".", collapse = "")
```

paste0 is the same as paste with an empty seperator
```{r}
paste0("X", 1:5)
```

NOTE: paste0 does not have a "sep" option that can be modified by the user

paste0("a", "b", sep = "c") == paste0("a", "b", "c")

paste0(rep(c("A","C","G","T"), each=4), c("A","C","G","T"), collapse = "")

### Extra material: Other string manipulations

Changing the case of strings
```{r}
string1 <- 'Data Science'
tolower(string1) 
toupper(string1)
```

### Exercise 1: Exercise with Strings

Create this string 'A&1B&2C&3' using a paste function
```{r}
paste(c("A","B","C"), 1:3, sep = "&", collapse = "")
```



## B. Working with Dates

```{r}
dates <- c('11/14/2011', '12/04/2012', '03/01/2013', '02/09/2019')
class(dates)
```

```{r}
real_dates <- as.Date(dates, format = '%m/%d/%Y')
class(real_dates)
```

```{r}
other_format <- format(real_dates, '%A %B %d, %Y')
class(other_format)
```


For the codes used to identify and format dates:
```{r}
?strptime
```

Identifying how long ago dates occurred
```{r}
today <- Sys.Date()

(dif <- today - real_dates)

class(dif)
```

To make a difference in times with a particular time unit of interest use difftime

```{r}
difftime(today, real_dates, units = "hours")
```


### Extra Materials: Lubridate

The lubridate package contains a powerful set of tools that can be used to extract and interact with dates.

There are functions like *mdy* that allow for simpler extration of date information from strings.
```{r}
install.packages("lubridate")
library(lubridate, quietly = TRUE)

(lubri_dates <- mdy(dates))
```

Extracting specific date information from a date object
```{r}
year(lubri_dates)

month(lubri_dates)

day(lubri_dates)
```


### Exercise 2: Exercise with Dates

a. Take the following date (November 11, 2011) and turn it into a date vector in R
b. Display the date vector in the format (month.day.year')

