---
title: "Language Diversity Dataset"
authors: 
  - name: Rashad Long
    email: RDLong718@gmail.com
  - name: Biyag Dukuray
    email: biyag15@gmail.com
format: html
editor: visual
---

## Original Dataset

**Data Source:** The Language Diversity Dataset, obtained from [untidydata repository](https://github.com/jvcasillas/untidydata)

**Justification for Reshaping:** The dataset is currently in an "untidy" format, meaning it is not in a wide format suitable for extensive analysis. Reshaping the data into a wide format would improve its usability and facilitate efficient analysis.

**Variable Renaming:** Additionally, certain variables within the dataset could be renamed for enhanced readability and clarity.

## Tidy Dataset Features

| Features         | Description                                                                                             | Data Type |
|------------------|---------------------------------------------------------------------------------------------------------|-----------|
| Continent        | Name of the Continent                                                                                   | chr       |
| Country          | Name of the Country                                                                                     | chr       |
| Languages        | Number of Languages Spoken                                                                              | num       |
| Area             | Area of land (km^2^)                                                                                    | num       |
| Population       | Population (Thousands)                                                                                  | num       |
| Weather Stations | The Number of weather stations used in calculation Mean Growth                                          | num       |
| Mean Growth      | The Mean Growing Season (months)                                                                        | num       |
| Growth Deviation | The standard deviation of the Growing Season values from the different weather stations in that country | num       |

: Variable Definintions

### Data Import

Dataset was automatically loaded into the environment directly from dev tools package. Can also be found in [repo.](https://github.com/jvcasillas/untidydata)

```{r view_data}
#| warning: FALSE
#| message: FALSE

library(untidydata)
library(tidyverse)
library(untidydata)

# Load the dataset
data(language_diversity)
str(language_diversity)
head(language_diversity)
```

### Data Tidying and Transformation

The data underwent a transformation to a wider format using the `pivot_wider()` function. This is done by using the names from the Measurement column and the values from the Value column. Subsequently, the column names were modified to enhance readability.

```{r tidy_data}
# Spread the data out using the pivot_wider function
wide_languages <- language_diversity |>
  pivot_wider(names_from = Measurement, values_from = Value)

# Change names of specific columns
wide_languages <- wide_languages |>
  rename(
    "Languages" = "Langs",
    "Weather Stations" = "Stations",
    "Mean Growth" = "MGS",
    "Growth Deviation" = "Std"
  )
str(wide_languages)
head(wide_languages)
```

### Analysis

This data will be analyzed to determine which countries have the highest amount of languages per capita. Some countries have a vast amount of different cultures and diversity and to determine which has the most we'll create a new column with the values of the amount of languages per capita. By calculating this average language count per individual, we aim to identify the countries within this dataset that boast the highest average number of languages spoken per resident.

```{r}
pivoted_data <-
  pivot_wider(language_diversity,
              names_from = Measurement,
              values_from = Value)

pivoted_data$Avg_Langs_Per_Person <-
  pivoted_data$Langs / pivoted_data$Population

print(pivoted_data)

```

### Conclusion

Based on the analysis of the dataset, the countries with the highest amount of languages per capita are Vanuatu, Papua New Guinea, French Guiana, Suriname and Gabon. The table above has been sorted by average amount of languages column which gives us this value.

```{r}
# | warning: FALSE

sorted_data <-
  pivoted_data[order(-pivoted_data$Avg_Langs_Per_Person),]

top_10 <- head(sorted_data, 10)

ggplot(top_10, aes(reorder(x = Country, Avg_Langs_Per_Person), y = Avg_Langs_Per_Person)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 Countries with the Highest Average Languages per Capita",
       x = "Country",
       y = "Average Languages per Capita")
```
